---
title: "2nd moment test"
author: "Thea Dowrich"
date: "10/17/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include = FALSE}
library(tidyverse)
library(splines)
```

```{r data, include = FALSE}
turbtrain <- read.csv("data-train.csv")
turbtrain$Sv <- turbtrain$St / turbtrain$Fr
```

```{r}
plot(turbtrain$R_moment_2)
plot(log(turbtrain$R_moment_2))
```



```{r check-ranges}
par(mfrow = c(2, 2))
hist(turbtrain$St)
hist(turbtrain$Re) #3 distinct groups
hist(turbtrain$Fr) #2 distinct groups (+ infinity)
hist(turbtrain$Sv)
```

```{r refactor}
turbtrain$Re <- as.factor(turbtrain$Re)
turbtrain$Fr <- as.factor(turbtrain$Fr)

# fr = St/ sv
# sv = st / fr
```


```{r relationships}
par(mfrow = c(1, 4))
plot(turbtrain$St, turbtrain$R_moment_2)
plot(turbtrain$Re, turbtrain$R_moment_2)
plot(turbtrain$Fr, turbtrain$R_moment_2)
plot(turbtrain$Sv, turbtrain$R_moment_2)
```
Most of the 2nd moment values for Re = 224 and Re = 398 are ~ 0, very small IQR (basically flat). Seeing similar results for `Fr`. 

- maybe transform the moment itself

```{r}
par(mfrow = c(1, 4))
plot(turbtrain$St, log(turbtrain$R_moment_2))
plot(turbtrain$Re, log(turbtrain$R_moment_2))
plot(turbtrain$Fr, log(turbtrain$R_moment_2))
plot(turbtrain$Sv, log(turbtrain$R_moment_2))
```
w log transformations, relationships are more apparent so might just stick with log(R_moment_2)

```{r regular-lm}
baselm <- lm(R_moment_2 ~ St + Re + Fr + Sv, data = turbtrain)
summary(baselm)
par(mfrow = c(2, 2))
plot(baselm)
```
p-value for St = 0.204 > 0.05, so potentially insignificant. But also Adj. R^2 = 0.4382, so hopefully log transformations improve it

```{r}
baselog <- lm(log(R_moment_2) ~ Re + Fr + St + Sv, data = turbtrain)
summary(baselog)
par(mfrow = c(2, 2))
plot(baselog)
```
Adj. R^2 = 0.7608 now, so big improvement!! p-value for Sv = 0.7226 >> 0.05, so will run a model with log transformation, but without Sv. 

```{r lm2}
lm2 <- lm(log(R_moment_2) ~ Re + Fr + St, data = turbtrain)
summary(lm2)
par(mfrow = c(2, 2))
plot(lm2)
```
Adj. R^2 = 0.7633, so improving even more and all predictors are significant. Will try with interactions next. 

```{r}
lm3 <- lm(R_moment_2 ~ Re*Fr + Re*St + Fr*St, data = turbtrain)
summary(lm3)
```
Tried with interaction terms and they do appear to be significant, so going to keep them in. Also, $R^2$ = 0.8631, so improving even more!!

```{r polynomials}
lm4 <- lm(R_moment_2 ~ Re*Fr + Re*St + Fr*St + poly(St, 5), data = turbtrain)
summary(lm4)
```
`St` only significant in quadratic form, so will keep using that and $R^2$ now 0.88 so improving even more.

```{r}
best_model <- lm(log(R_moment_2) ~ Re*Fr + Re*St + Fr*St + poly(St, 2), data = turbtrain)
summary(best_model)
par(mfrow = c(2,2))
plot(lm5)
```







