---
title: "Turbulence Case Study"
author: "Alexander Bendeck, Anna Darwish, Thea Dowrich, Ashley Murray, Matty Pahren"
geometry: "left=1.25cm,right=1.25cm,top=1.3cm,bottom=1.3cm"
date: "10/19/2020"
output: pdf_document
---

```{r, function checking for installed packages, echo=FALSE, results="hide", include=FALSE}
# Validate that all necessary packaged have been downloaded, install otherwise or throw err package DNE
pkgTest <- function(x)
{
  if (!require(x,character.only = TRUE))
  {
    install.packages(x,repos = "http://cran.r-project.org", dep=TRUE)
    if(!require(x,character.only = TRUE)) stop("Package not found")
  }
}
```

```{r Package Test Libraries, echo=FALSE, results="hide", include=FALSE}
pkgTest("boot")
pkgTest("dplyr")
pkgTest("ggplot2")
pkgTest("tidyr")
pkgTest("knitr")
pkgTest("kableExtra")
pkgTest("gridExtra")
pkgTest("pander")
pkgTest("tidyverse")
```

```{r Load in Libraries, echo=FALSE, results="hide", include=FALSE}
library(boot)
library(dplyr)
library(ggplot2)
library(tidyr)
library(knitr)
library(kableExtra)
library(gridExtra)
library(pander)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
ggplot2::theme_set(new = theme_bw())
set.seed(123)
```

# Introduction

One of the greatest areas of research in Physics today is the study of fluid motion [[1]][Bibliography]. The applications of these investigations span across the aviation industry, astrophysics, oil spills in oceans, and more [[1]][Bibliography]. Turbulence is defined to be "chaotic changes in pressure and flow velocity" [[2]][Bibliography]. For this case study, we will be exploring models to better understand and predict particle volume cluster under homogeneous isotropic turbulence.

Our covariates of concern in this study are *Reynolds number (Re)*, *Froude's Number (Fr)*, and *Particle characteristic (St)*. *Reynolds number* captures the ratio between fluid velocity and the body size of the object moving through the fluid [[1]][Bibliography]. *Froude's Number* specifies the gravity conditions of the fluid, and *Particle characteristic* provides information on the size and density of the particles [[1]][Bibliography]. We will be modeling these covariates' relationship with the first four raw moments of particle cluster volumes in an idealized turbulence. Please refer to the Data Dictionary in the Appendix for more details on these variables.

For this case study, our research objectives are:

1. Understand and interpret the relationship between the first four moments of particle cluster volumes and *Reynolds number*, *Froude's Number*, and *Particle characteristic*.
2. Estimate the accuracy of our models.

# Methodology

Prior studies on turbulence, as well as the presentation in lab by Reza and Jonathan, motivated us to consider the covariate of *Settling parameter (Sv)*, which is the ratio of *Stokes number* to *Froude's Number*. This covariate is continuous and has finite values, so it was used in place of *Froude's number*, which had only three levels. Further, we chose to treat *Reynold's number* as a categorical covariate with three levels: Low, Med., and High. While this may lead to ambiguity in generalizing this model outside of the scope of these training and test sets, both datasets were constrained to these three categories, making it difficult to quantify its linear relationship with the four moments.

For each of our four models, we began by conducting exploratory data analysis with the predictors in our dataset. This informed our decisions for possible models, transformations, and interactions to explore. Given that one of our primary goals is to understand and interpret the relationship between the first four moments of particle volume cluster and our predictors, we chose to use linear regression models. We also explored polynomial and spline models to determine whether either provided significantly better predictive power.

For the second, third, and fourth moment, we noticed a curved relationship between the response and covariates, which motivated us to perform a logged transformation on the response.

To compare different linear regression models for the same moment, we began by comparing adjusted $R^2$ values, as it compensates for added predictors that do not significantly improve the amount of explained variance in the response. To compare any two models across the different methodologies we attempted, we performed 5-fold cross validation and compared the estimated test errors.

Finally, we examined residual and leverage plots for each of our four models to ensure we didn't violate any of the assumptions that accompany linear regression models. All residual and leverage plots are located in the Appendix.

# Results and Discussion

## Moment 1

```{r Read in Train Data, echo=FALSE}
turbulence = read.csv("data-train.csv")

turbulence$Sv = turbulence$St / turbulence$Fr

turbulence$Re <- as.factor(turbulence$Re)
levels(turbulence$Re) <- c("Low", "Med", "High")

turbulence$Fr <- as.factor(turbulence$Fr)
levels(turbulence$Fr) <- c("Low", "Inf", "Med") 

turbulence$Sv_2 = turbulence$Sv^2
turbulence$St_2 = turbulence$St^2
```

```{r Read in Test Data, echo=FALSE}
test = read.csv("data-test.csv")
test$Sv = test$St / test$Fr

test$Re <- as.factor(test$Re)
levels(test$Re) <- c("Low", "Med", "High")

test$Fr <- as.factor(test$Fr)
levels(test$Fr) <- c("Low", "Inf", "Med") 

test$St_2 = test$St^2
test$Sv_2 = test$Sv^2
```

```{r Moment 1 Model, fig.align="center", fig.width=6, fig.height =4}
moment.1.mod <- lm(R_moment_1 ~ Re + St + St * Re + Sv + Sv * Re + Sv_2 + Sv_2 * Re, data=turbulence)
moment.1.mod.glm <- glm(R_moment_1 ~ Re + St + St*Re + Sv + Sv*Re + Sv_2 + Sv_2*Re, data=turbulence)
pander(moment.1.mod, caption="Moment 1")

m1.preds.raw = data.frame(predict(moment.1.mod, newdata=test, interval="prediction", level=0.95))
m1.preds = m1.preds.raw %>% mutate(lwr = case_when(lwr <= 0 ~ 0,
                                                   TRUE ~ lwr))
full_preds_df = cbind(test, m1.preds)

full_preds_df = full_preds_df %>% dplyr::select(c("Re", "St", "Sv", "fit", "lwr", "upr"))

colnames(full_preds_df) = c("Re", "St", "Sv", "pred", "95% low pred", "95% high pred")

preds_of_interest = full_preds_df[c(3, 10, 14, 15),]
rownames(preds_of_interest) = NULL

pander(preds_of_interest, caption="Selected Predictions for Moment 1")
```

From the model output, we see that for the baseline level (low Reynolds number), increased Stokes number and settling parameter have a notable effect of increasing the first moment. Settling parameter has a nonlinear effect. Having a medium or high Reynolds number effectively lowers the intercept in these cases to 0, since the main effects for these two factor levels basically cancel out the overall intercept term. Similarly, the interaction between Stokes number and both medium and high Reynolds number seems to cancel out the main effect of Stokes number. In other words, the effect of increasing Stokes number on the first moment is close to 0 when the Reynolds number is medium or high. This is likewise true for the effect of settling parameter.

It makes sense that a medium or high Reynolds number greatly reduces the first moment. The first moment represents the average size of particle clusters. The Reynolds number measures turbulence, so if the turbulence is moderate to high, we would expect to not see very large particle clusters due to turbulent flow. It also makes sense that increasing the Stokes number, a measure of inertia, would cause greater cluster sizes, but only when there is low turbulence.

Consistent with the discussion above, we see that when predicting the first moment on new data, the predictions are very close to 0 when the Reynolds number is high or medium. We also see that in these cases, the upper 95% confidence interval leaves quite a bit of uncertainty around the prediction (note that the lower bounds of these CIs have been truncated at 0). However, when predicting the first moment when the Reynolds number is low, we predict higher first moments overall compared to when the Reynolds number is high or medium. We also see that increasing Stokes number and settling parameter increase the first moment prediction notably when the Reynolds number is low. Also, since the predictions are further from 0 with a low Reynolds number, the width of the confidence intervals is small compared to the predictions themselves.

We evaluated our assumptions associated with this model and observed a roughly even scatter about the 0 line in the residual plot, along with generally low Cook's Distance values. 5-fold cross-validation yielded an average test error of `r round(cv.glm(turbulence, moment.1.mod.glm)$delta[1],5)` 

## Moment 2

```{r Moment 2 Model, fig.align="center", fig.width=8, fig.height =4}
og.mod.2 <- lm(log(R_moment_2) ~ Re*Fr + Re*St + poly(St, 2), data = turbulence) # This is a rank deficient fit so the estimates of the coefficients are unreliable
moment.2.mod.glm <- glm(log(R_moment_2) ~ Re + Sv + St + St_2 + Sv*Re, data = turbulence)
moment.2.mod <- lm(log(R_moment_2) ~ Re + Sv + St + St_2 + Sv*Re, data = turbulence)
pander(moment.2.mod, caption="Moment 2")

m2.preds.raw = data.frame(predict(moment.2.mod, newdata=test, interval="prediction", level=0.95))
m2_full_preds_df = cbind(test, m2.preds.raw)
m2_full_preds_df = m2_full_preds_df %>% dplyr::select(c("Re", "St", "Sv", "Sv_2", "fit", "lwr", "upr"))
colnames(m2_full_preds_df) = c("Re", "Sv", "St","St_2", "pred", "95% low pred", "95% high pred")
m2_preds_of_interest = m2_full_preds_df[c(3, 10, 14, 15),]
rownames(m2_preds_of_interest) = NULL
pander(m2_preds_of_interest, caption="Selected Predictions for Moment 2")
```

From the model output, we see that as the Reynold's number increases, we expect the log of the second moment of our response to decrease. This suggests there is less variance in particle cluster volumes as the ratio between fluid velocity and body size increases. This effect cannot be understood in isolation however, because of the interaction effect with the settling parameter, *Sv*. Despite the main effect of the settling parameter suggesting an increase in the log of the second moment of the response, this effect is reversed for higher levels of Reynold's number. Finally, there appears to be a quadratic relationship between the log of the second moment and the Stoke's number.

From 5-fold cross validation, we achieve an estimated test error of `r round(cv.glm(turbulence, moment.2.mod.glm)$delta[1],3)`. From the abbreviated table of predictions, we see that our uncertainty worsens for larger values of Stoke's number. These predictions reflect our model's results, in that increasing Reynold's number generally lends to a lower estimate of the log of  the second moment, even if the settling parameter values increases.

## Moment 3

```{r Moment 3 Model, fig.align="center", fig.width=8, fig.height =4}
og.moment.3.mod <- glm(log(R_moment_3) ~ Fr + St + Re + Fr*St + Fr*Re + St*Re, data=turbulence) # also rank deficient fit
moment.3.mod.glm <- glm(log(R_moment_3) ~ Sv_2 + Sv + Re, data=turbulence)
moment.3.mod <- lm(log(R_moment_3) ~ Sv_2 + Sv + Re, data=turbulence)
pander(moment.3.mod, caption="Moment 3")

m3.preds.raw = data.frame(predict(moment.3.mod, newdata=test, interval="prediction", level=0.95))
m3_full_preds_df = cbind(test, m3.preds.raw)
m3_full_preds_df = m3_full_preds_df %>% dplyr::select(c("Re", "Sv", "Sv_2", "fit", "lwr", "upr"))
colnames(m3_full_preds_df) = c("Re", "Sv", "Sv_2", "pred", "95% low pred", "95% high pred")
m3_preds_of_interest = m3_full_preds_df[c(3, 10, 14, 15),]
rownames(m3_preds_of_interest) = NULL
pander(m3_preds_of_interest, caption="Selected Predictions for Moment 3")
```

For this model, interaction effects lead to higher estimated test errors and were not significant in terms of their p-values. Therefore, in our final model, there are no interaction effects for predicting the log of the 3rd moment. From the model output, we see that as the Reynold's number increases, we expect the log of the third moment of our response to decrease. This suggests the "lopsidedness" of the distribution of the log of the 3rd moment decreases as the ratio between fluid velocity and body size increases. Finally, there appears to be an increasing relationship in the log of the third moment and the *linear* effect of the settling parameter and a decreasing relationship with the *quadratic* effect of the settling parameter. This suggests that there is a leveling off of the effect of the settling parameter.

From 5-fold cross validation, we achieve an estimated test error of `r round(cv.glm(turbulence, moment.3.mod.glm)$delta[1],3)`. From the abbreviated table of predictions, we see that our uncertainty slightly worsens for higher values of Reynold's number. These predictions reflect our model's results in that increasing Reynold's number generally lends to a lower estimate of the log of the third moment and generally increases for larger values of the settling parameter.

## Moment 4

```{r Moment 4 Model, fig.align="center", fig.width=8, fig.height =4}
moment.4.mod <- lm(log(R_moment_4) ~ Re + Sv + St + Re*Sv + Re*St + Sv*St, data=turbulence)
moment.4.mod.glm <- glm(log(R_moment_4) ~ Re + Sv + St + Re*Sv + Re*St + Sv*St, data=turbulence)
pander(moment.4.mod, caption="Moment 4")
```

Holding all else constant, we expect the log of the fourth moment for observations with a medium or high Reynolds number to be lower than that of observations with a low Reynolds number. For the baseline level (a low Reynolds number), increasing the settling parameter also increases the log of the fourth moment. However, for medium and high Reynolds numbers, increasing the settling parameter does not lead to as large of an increase in the log of the fourth moment, since the original settling parameter coefficient is reduced by the interactions. From 5-fold cross-validation we achieved an estimated test error of `r round(cv.glm(turbulence, moment.4.mod.glm)$delta[1],3)`.

```{r moment-4-pred}
m4.preds.raw = data.frame(predict(moment.4.mod, newdata=test, interval="prediction", level=0.95))
m4_full_preds_df = cbind(test, m4.preds.raw)
m4_full_preds_df = m4_full_preds_df %>% dplyr::select(c("Re", "St", "Sv", "fit", "lwr", "upr"))
colnames(m4_full_preds_df) = c("Re", "St", "Sv", "pred", "95% low pred", "95% high pred")
m4_preds_of_interest = m4_full_preds_df[c(6, 21, 4, 14),]
rownames(m4_preds_of_interest) = NULL
pander(m4_preds_of_interest, caption="Selected Predictions for Moment 4")
```

Consistent with what we saw above, for constant settling parameter and Stokes numbers, observations with high Reynolds numbers have lower log fourth moments when compared to observations with low Reynolds number. Additionally, the higher settling parameter led to an even higher prediction for the observation with a low Reynolds number compared to the observation with a high Reynolds number. However, it is worth noting that these predictions have very wide intervals, indicating that it would be possible to see opposite scenarios occur in nature. 

# Conclusion

Our goals in this case study were to 1) model and interpret the relationship between *Reynolds number (Re)*, *Froude's Number (Fr)*, *Particle characteristic (St)*, and the *Settling Parameter (Sv)* and the first four moments of particle cluster volumes and 2) estimate the test errors of our models. We accomplished this by first exploring various model types, interaction effects, and transformations. After, we chose our final models with cross validation and variable selection methods. One of the key findings of our models is that the *Settling Parameter* served as a stronger predictor in place of *Froude's Number (Fr)* across all four models. Further, log transforming the the second, third, and fourth moments of particle cluster volumes was necessary to estimate a linear relationship with our covariates. Finally, for all moments except the third moment, the relationship between particle volume cluster and the *Reynolds number (Re)* is dependent upon the value of the *Settling Parameter (Sv)*. For future studies, future models may wish to focus more directly on prediction to achieve a lower estimated test error, along with other values of the *Reynolds number (Re)* and *Froude's number (Fr)*.

\newpage

# Appendix

## Data Dictionary

```{r Data Dictionary, echo=F}
data.dict <- data.frame(Variable = c(rep("Reynolds number (Re)", 1),
                                          rep("Froude's Number (Fr)", 1),
                                          rep("Particle characteristic (St)", 1),
                                          rep("Settling Parameter (Sv)", 1)),
                               Values = linebreak(c("90 (Low), 224 (Med), 398 (High)",
                                                    "0.052 (Low), 0.300 (Med), Inf (High)",
                                                    "[0.05, 3.00]",
                                                    "[0.00,57.692]")))
kable(data.dict, escape=F, booktabs = T, align = "ll", caption = "Data Dictionary") %>%
  kable_styling(latex_options = c("hold_position")) %>%
  column_spec(1, bold=T) %>%
  collapse_rows(columns = 1, latex_hline = "major", valign = "middle")
```

## Residual and Leverage Plots

```{r fig.height=3,fig.width=8}
par(mfrow = c(1, 2))
plot(moment.1.mod, which=c(1))
plot(moment.1.mod, which=c(5))
```

```{r fig.height=3,fig.width=8}
par(mfrow = c(1, 2))
plot(moment.2.mod, which=c(1))
plot(moment.2.mod, which=c(5))
```

```{r fig.height=3,fig.width=8}
par(mfrow = c(1, 2))
plot(moment.3.mod, which=c(1))
plot(moment.3.mod, which=c(5))
```

```{r fig.height=3,fig.width=8}
par(mfrow = c(1, 2))
plot(moment.4.mod, which=c(1))
plot(moment.4.mod, which=c(5))
```

# Bibliography

1. Momenifar, R., &amp; Holt, J. (2020). STA 325 - Lab 1. Lecture.
2. Turbulence. (2020, October 12). Retrieved October 19, 2020, from https://en.wikipedia.org/wiki/Turbulence
3. Moment (Mathematics). (2020 October 20). Retrieved October 24, 2020 from https://en.wikipedia.org/wiki/Moment_(mathematics)#:~:text=The%20third%20central%20moment%20is,called%20the%20skewness%2C%20often%20%CE%B3.




